The role of prenatal conditions on brain mitochondrial physiology and prey detection in a lizard

### 2025-04-01 ms.qmd
updates

#### How are mitochondrial physiology and oxidative stress related to prey detection?
```{r, model_sem_OB}
#| label: model_sem_OB
# Making the SEM model by using a multivariate brms. The aim is to test the relationships between mitochondrial physiology and detection latency.
# To simplify the random factors, and since there is no effect of experience with the prey in prey detection, latency was averaged across the two preys. Every other predictor added to the model was based on previous brms for each variable separated.
# All continue variables were standardized (var/2SD) before running the models (see data_process.R). 
#
source(here("R", "func.R"))
#
SEM_df <- clean_df %>%
  group_by(lizard_id, prey) %>%
  mutate(t_D = mean(t_D, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(prey == "Unknown")
#
#
data_sem_OB <- SEM_df %>%
  filter(region == "OB") %>%
  mutate(obs = as.integer(c(1:80))) %>%
  mutate(vec = rep(1, length(obs)))
# Create the models
#
refit <- FALSE
#
#
if(refit){
  m_OB <- brm(
    bf(t_D | se(vec, sigma = FALSE) ~ cort + mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch) + (1|p|obs)) +
    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ age_euthanasia + mean_ros + (1|clutch)+ (1|p|obs)) +
    bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
  set_rescor(FALSE),
  family = gaussian(),
  data = data_sem_OB,
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OB, file = here("output/m_SEM/m_OB.rds"))
} else {
  m_OB <- readRDS(here("output/m_SEM/m_OB.rds"))
}
```

@@

OB_coeff_cort_det <- post_sem_OB$b_tD_cortCORT
det_sem_OB_control <- post_sem_OB %>%
  dplyr::select(-b_tD_cortCORT) %>%
  dplyr::sample_n(size = 12000, replace = FALSE) # We take here 12000 random values for all the variables to get estimates for Control treatment, which is the reference level
det_sem_OB_cort <- post_sem_OB %>%
  mutate(across(everything(), ~.x + b_tD_cortCORT)) %>%
  dplyr::sample_n(size = 12000, replace = FALSE) %>% # We take here 12000 random values for all the variables AFTER adding to b_tD_cortCORT to all columns to get estimates for CORT lizards, which is NOT the reference level
  dplyr::select(-b_tD_cortCORT)
det_sem_OB_controlled <- bind_rows(det_sem_OB_control, det_sem_OB_cort)
OB_coeff_mitodensity_det <- det_sem_OB_controlled$b_tD_mean_mitodensity
OB_coeff_potential_det <- det_sem_OB_controlled$b_tD_mean_potential
OB_coeff_dna_det <- det_sem_OB_controlled$b_tD_mean_dnadamage
OB_coeff_perox_det <- det_sem_OB_controlled$b_tD_mean_peroxidation

@@

OB_coeff_age_dna <- post_sem_OB$b_meandnadamage_age_euthanasia

@@

OB_coeff_age_perox <- post_sem_OB$b_meanperoxidation_age_euthanasia

@@

OB_undir_age_det <- OB_coeff_age_dna * OB_coeff_dna_det + OB_coeff_age_perox * OB_coeff_perox_det

@@

OB_total_cort_det <- OB_coeff_cort_det

@@

OB_total_age_det <- OB_undir_age_det

@@

OB_total_age_dna <- OB_coeff_age_dna

@@

OB_total_age_perox <- OB_coeff_age_perox

@@

  reference_variable = rep("t_D", 7),
  predictor_modulator = c("cort",
                          "age",
                          "mean_mitodensity",

@@

  direct_effects = I(list(OB_coeff_cort_det,
                          OB_coeff_mitodensity_det,

@@

                          OB_undir_age_det,
                          OB_undir_mitodensity_det,

@@

  total_effects = I(list(OB_total_cort_det,
                        OB_total_age_det,
                        OB_total_mitodensity_det,

@@

  reference_variable = rep("DNA damage", 4),
  predictor_modulator = c("age",
                          "mean_mitodensity",

@@

  direct_effects = I(list(OB_coeff_age_dna,

@@

                          OB_undir_mitodensity_dna,

@@

  total_effects = I(list(OB_total_age_dna,
                        OB_total_mitodensity_dna,

@@

  reference_variable = rep("lipid peroxidation", 4),
  predictor_modulator = c("age",
                          "mean_mitodensity",

@@

  direct_effects = I(list(OB_coeff_age_perox,

@@

                          OB_undir_mitodensity_perox,

@@

  total_effects = I(list(OB_total_age_perox,
                        OB_total_mitodensity_perox,

@@

```{r, fig-sem_results_OB}
#| label: fig-sem_results_OB
#| fig-cap: "Structural Equation Models for OB/Chemical stimulus"
#| fig-name: "fig-sem_results_OB"

@@

# A) Getting all the direct coefficients
OB_cort_det <- paste0(format_dec(mean(OB_coeff_cort_det), 3),
                  " [", format_dec(quantile(OB_coeff_cort_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_cort_det, 0.95), 3), "]")
OB_density_det <- paste0(format_dec(mean(OB_coeff_mitodensity_det), 3),
                  " [", format_dec(quantile(OB_coeff_mitodensity_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_mitodensity_det, 0.95), 3), "]")
OB_potential_det <- paste0(format_dec(mean(OB_coeff_potential_det), 3),
                  " [", format_dec(quantile(OB_coeff_potential_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_potential_det, 0.95), 3), "]")
OB_dna_det <- paste0(format_dec(mean(OB_coeff_dna_det), 3),
                  " [", format_dec(quantile(OB_coeff_dna_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_dna_det, 0.95), 3), "]")
OB_perox_det <- paste0(format_dec(mean(OB_coeff_perox_det), 3),
                  " [", format_dec(quantile(OB_coeff_perox_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_perox_det, 0.95), 3), "]")

@@

OB_age_dna <- paste0(format_dec(mean(OB_coeff_age_dna), 3),
                  " [", format_dec(quantile(OB_coeff_age_dna, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_age_dna, 0.95), 3), "]")
OB_ros_dna <- paste0(format_dec(mean(OB_coeff_ros_dna), 3),
                  " [", format_dec(quantile(OB_coeff_ros_dna, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_ros_dna, 0.95), 3), "]")

@@

OB_age_perox <- paste0(format_dec(mean(OB_coeff_age_perox), 3),
                  " [", format_dec(quantile(OB_coeff_age_perox, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_age_perox, 0.95), 3), "]")
OB_ros_perox <- paste0(format_dec(mean(OB_coeff_ros_perox), 3),
                  " [", format_dec(quantile(OB_coeff_ros_perox, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_ros_perox, 0.95), 3), "]")

@@

OB_density_ros <- paste0(format_dec(mean(OB_coeff_mitodensity_ros), 3),
                  " [", format_dec(quantile(OB_coeff_mitodensity_ros, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_mitodensity_ros, 0.95), 3), "]")
OB_potential_ros <- paste0(format_dec(mean(OB_coeff_potential_ros), 3),
                  " [", format_dec(quantile(OB_coeff_potential_ros, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_potential_ros, 0.95), 3), "]")

@@

imgOB <- readPNG(here("Others", "SEM_OB.png"))
plot_SEM_OB <- rasterGrob(imgOB, interpolate = TRUE)
fig_SEM_OB <- ggdraw(plot_SEM_OB) +
  annotate("text", x = 0.815, y = 0.59, label = OB_cort_det, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.35, y = 0.26, label = OB_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.35, y = 0.918, label = OB_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.745, label = OB_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.455, label = OB_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.55, y = 0.538, label = OB_age_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.525, y = 0.813, label = OB_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.57, y = 0.115, label = OB_age_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times", color = "darkgreen") +
  annotate("text", x = 0.525, y = 0.365, label = OB_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.285, y = 0.355, label = OB_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.285, y = 0.825, label = OB_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
ggsave(here("./output/figures/text/SEM_OB.png"), plot = fig_SEM_OB, width = 21, height = 10, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/SEM_OB.png")

```{r, models_sem_OT}
#| label: models_sem_OT
# Making the same models above but for OT/Visual stimulus
#
source(here("R", "func.R"))
#
#
data_sem_OT <- SEM_df %>%
  filter(region == "OT") %>%
  mutate(obs = as.integer(c(1:80))) %>%
  mutate(vec = rep(1, length(obs)))
#
# Create the models
#
refit <- FALSE
#
if(refit){
  m_OT <- brm(
    bf(t_D | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|p|obs)) +
    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ cort*temp + sex + mean_ros + (1|clutch) + (1|q|obs)) +
    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ cort*temp + age_euthanasia + mean_ros + (1|clutch) + (1|q|obs)) +
    bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
  set_rescor(FALSE),
  family = gaussian(),
  data = data_sem_OT, 
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OT, file = here("output/m_SEM/m_OT.rds"))
} else {
  m_OT <- readRDS(here("output/m_SEM/m_OT.rds"))
}
```

```{r, sem_tidy_OT}
#| label: sem_tidy_OT
source(here("R", "func.R"))
#
# I am extracting here all the values for getting the total effects of each of the variables in the model. I am using the posterior values for each of the variables to get the total effects assuming that:
## total effect = direct effect + indirect effect + residual correlation
# In other words:
## total effect = 
# Extract the posteriors for the SEM model
post_sem_OT <- as_draws_df(m_OT) 
#
#### A) Get the direct paths per each variable
# 
# A.1) Detection
OT_coeff_mitodensity_det <- post_sem_OT$b_tD_mean_mitodensity
OT_coeff_potential_det <- post_sem_OT$b_tD_mean_potential
OT_coeff_dna_det <- post_sem_OT$b_tD_mean_dnadamage
OT_coeff_perox_det <- post_sem_OT$b_tD_mean_peroxidation
#
# A.2) DNA damage
#
# Getting CORT coefficient for both temperatures controlling for each sex
dna_sem_OT_male <- post_sem_OT %>%
  dplyr::select(-b_meandnadamage_sexFemale)
dna_sem_OT_fem <- post_sem_OT %>%
  mutate(across(everything(), ~.x + b_meandnadamage_sexFemale)) %>%
  dplyr::select(-b_meandnadamage_sexFemale) # Add effect of sex to everything to see estimates on females (males the reference level for sex)

dna_sem_OT_male_trt <- dna_sem_OT_male %>%
  dplyr::sample_n(size = 12000, replace = FALSE)
dna_sem_OT_fem_trt <- dna_sem_OT_fem %>%
  dplyr::sample_n(size = 12000, replace = FALSE)

dna_sem_OT_trt <- bind_rows(dna_sem_OT_male_trt, dna_sem_OT_fem_trt)

OT_coeff_cort_dna_cold <- dna_sem_OT_trt$b_meandnadamage_cortCORT
OT_coeff_cort_dna_hot <- dna_sem_OT_trt$b_meandnadamage_Intercept + dna_sem_OT_trt$b_meandnadamage_cortCORT + dna_sem_OT_trt$b_meandnadamage_tempHot + dna_sem_OT_trt$`b_meandnadamage_cortCORT:tempHot` - (dna_sem_OT_trt$b_meandnadamage_Intercept + dna_sem_OT_trt$b_meandnadamage_tempHot)
# Getting ros coefficient controlling for each level of the cort-temp interaction and controlling for each sex
dna_sem_OT_male_controlcold <- dna_sem_OT_male %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_male_controlhot <- dna_sem_OT_male %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_male_cortcold <- dna_sem_OT_male %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_male_corthot <- dna_sem_OT_male %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
                `b_meandnadamage_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)

dna_sem_OT_fem_controlcold <- dna_sem_OT_fem %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_fem_controlhot <- dna_sem_OT_fem %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_fem_cortcold <- dna_sem_OT_fem %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)
dna_sem_OT_fem_corthot <- dna_sem_OT_fem %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
                `b_meandnadamage_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 3000, replace = FALSE)

dna_sem_OT_controlled <- bind_rows(dna_sem_OT_male_cortcold,
                                  dna_sem_OT_male_corthot,
                                  dna_sem_OT_male_controlcold,
                                  dna_sem_OT_male_controlhot,
                                  dna_sem_OT_fem_cortcold,
                                  dna_sem_OT_fem_corthot,
                                  dna_sem_OT_fem_controlcold,
                                  dna_sem_OT_fem_controlhot)

OT_coeff_ros_dna <- dna_sem_OT_controlled$b_meandnadamage_mean_ros
#
# Getting sex coefficient controlling for each level of the cort-temp interaction
dna_sem_OT_controlcold <- post_sem_OT %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
dna_sem_OT_cortcold <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
dna_sem_OT_controlhot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_tempHot)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
dna_sem_OT_corthot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meandnadamage_cortCORT + b_meandnadamage_tempHot +
                `b_meandnadamage_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)

dna_sem_OT_sex <- bind_rows(dna_sem_OT_cortcold,
                            dna_sem_OT_controlcold,
                            dna_sem_OT_corthot,
                            dna_sem_OT_controlhot)

OT_coeff_sex_dna <- dna_sem_OT_sex$b_meandnadamage_sexFemale
#
# A.3) Lipid peroxidation
OT_coeff_cort_perox_cold <- post_sem_OT$b_meanperoxidation_cortCORT
OT_coeff_cort_perox_hot <- post_sem_OT$b_meanperoxidation_Intercept + post_sem_OT$b_meanperoxidation_cortCORT + post_sem_OT$b_meanperoxidation_tempHot + post_sem_OT$`b_meanperoxidation_cortCORT:tempHot` - (post_sem_OT$b_meanperoxidation_Intercept + post_sem_OT$b_meanperoxidation_tempHot)

# Getting the coefficient of age and ros controlling for each level of the cort-temp interaction
perox_sem_OT_controlcold <- post_sem_OT %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
perox_sem_OT_cortcold <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_cortCORT)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
perox_sem_OT_controlhot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_tempHot)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)
perox_sem_OT_corthot <- post_sem_OT %>%
  dplyr::mutate(across(everything(), ~.x + b_meanperoxidation_cortCORT + b_meanperoxidation_tempHot +
                `b_meanperoxidation_cortCORT:tempHot`)) %>%
  dplyr::sample_n(size = 6000, replace = FALSE)

perox_sem_OT_controlled <- bind_rows(perox_sem_OT_cortcold,
                            perox_sem_OT_controlcold,
                            perox_sem_OT_corthot,
                            perox_sem_OT_controlhot)

OT_coeff_ros_perox <- perox_sem_OT_controlled$b_meanperoxidation_mean_ros
OT_coeff_age_perox <- perox_sem_OT_controlled$b_meanperoxidation_age_euthanasia

@@

OT_undir_cort_det_COLD <- OT_coeff_cort_dna_cold * OT_coeff_dna_det + OT_coeff_cort_perox_cold * OT_coeff_perox_det
OT_undir_cort_det_HOT <- OT_coeff_cort_dna_hot * OT_coeff_dna_det + OT_coeff_cort_perox_hot * OT_coeff_perox_det
OT_undir_age_det <- OT_coeff_age_perox * OT_coeff_perox_det
OT_undir_sex_det <- OT_coeff_sex_dna * OT_coeff_dna_det

@@

OT_total_cort_det_COLD <- OT_undir_cort_det_COLD
OT_total_cort_det_HOT <- OT_undir_cort_det_HOT
OT_total_age_det <- OT_undir_age_det
OT_total_sex_det <- OT_undir_sex_det

@@

OT_total_cort_dna_COLD <- OT_coeff_cort_dna_cold
OT_total_cort_dna_HOT <- OT_coeff_cort_dna_hot
OT_total_sex_dna <- OT_coeff_sex_dna

@@

OT_total_cort_perox_COLD <- OT_coeff_cort_perox_cold
OT_total_cort_perox_HOT <- OT_coeff_cort_perox_hot
OT_total_age_perox <- OT_coeff_age_perox

@@

  reference_variable = rep("t_D", 9),
  predictor_modulator = c("cort_cold",
                          "cort_hot",
                          "age",
                          "sex",
                          "mean_mitodensity",

@@

                          OT_coeff_mitodensity_det,

@@

  indirect_effects = I(list(OT_undir_cort_det_COLD,
                          OT_undir_cort_det_HOT,
                          OT_undir_age_det,
                          OT_undir_sex_det,
                          OT_undir_mitodensity_det,

@@

  total_effects = I(list(OT_total_cort_det_COLD,
                        OT_total_cort_det_HOT,
                        OT_total_age_det,
                        OT_total_sex_det,
                        OT_total_mitodensity_det,

@@

  reference_variable = rep("DNA damage", 6),
  predictor_modulator = c("cort_cold",
                          "cort_hot",
                          "sex",
                          "mean_mitodensity",

@@

  direct_effects = I(list(OT_coeff_cort_dna_cold,
                          OT_coeff_cort_dna_hot,
                          OT_coeff_sex_dna,

@@

                          OT_undir_mitodensity_dna,

@@

  total_effects = I(list(OT_total_cort_dna_COLD,
                        OT_total_cort_dna_HOT,
                        OT_total_sex_dna,
                        OT_total_mitodensity_dna,

@@

  reference_variable = rep("Lipid peroxidation", 6),
  predictor_modulator = c("cort_cold",
                          "cort_hot",
                          "age",
                          "mean_mitodensity",

@@

  direct_effects = I(list(OT_coeff_cort_perox_cold,
                          OT_coeff_cort_perox_hot,
                          OT_coeff_age_perox,

@@

                          OT_undir_mitodensity_perox,

@@

  total_effects = I(list(OT_total_cort_perox_COLD,
                        OT_total_cort_perox_HOT,
                        OT_total_age_perox,
                        OT_total_mitodensity_perox,

@@

#### Results of the multivariate models

>>>>

#### How are mitochondrial physiology and oxidative stress related to prey detection?
```{r, model_sem_OB}
#| label: model_sem_OB
# Making the SEM model by using a multivariate brms. The aim is to test the relationships between mitochondrial physiology and detection latency.
# To simplify the random factors, and since there is no effect of experience with the prey in prey detection, latency was averaged across the two preys. Every other predictor added to the model was based on previous brms for each variable separated.
# All continue variables were standardized (var/2SD) before running the models (see data_process.R). 
#
source(here("R", "func.R"))
#
SEM_df <- clean_df %>%
  group_by(lizard_id, prey) %>%
  mutate(t_D = mean(t_D, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(prey == "Unknown")
#
#
data_sem_OB <- SEM_df %>%
  filter(region == "OB") %>%
  mutate(obs = as.integer(c(1:80))) %>%
  mutate(vec = rep(1, length(obs)))
# Create the models
#
refit <- FALSE
#
#
if(refit){
  m_OB <- brm(
    bf(t_D | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|q|obs)) +
    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch) + (1|p|obs)) +
    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch)+ (1|p|obs)) +
    bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
  set_rescor(FALSE),
  family = gaussian(),
  data = data_sem_OB,
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OB, file = here("output/m_SEM/m_OB.rds"))
} else {
  m_OB <- readRDS(here("output/m_SEM/m_OB.rds"))
}
```

@@

OB_coeff_mitodensity_det <- post_sem_OB$b_tD_mean_mitodensity
OB_coeff_potential_det <- post_sem_OB$b_tD_mean_potential
OB_coeff_dna_det <- post_sem_OB$b_tD_mean_dnadamage
OB_coeff_perox_det <- post_sem_OB$b_tD_mean_peroxidation

@@

@@

@@

@@

@@

@@

@@

@@

  reference_variable = rep("t_D", 5),

@@

  direct_effects = I(list(OB_coeff_mitodensity_det,

@@

  indirect_effects = I(list(OB_undir_mitodensity_det,

@@

  total_effects = I(list(OB_total_mitodensity_det,

@@

  reference_variable = rep("DNA damage", 3),

@@

@@

  indirect_effects = I(list(OB_undir_mitodensity_dna,

@@

  total_effects = I(list(OB_total_mitodensity_dna,

@@

  reference_variable = rep("lipid peroxidation", 3),

@@

@@

  indirect_effects = I(list(OB_undir_mitodensity_perox,

@@

  total_effects = I(list(OB_total_mitodensity_perox,

@@

```{r, fig-sem_results_OB}
#| label: fig-sem_results_OB
#| fig-cap: "Structural Equation Models for OB/Chemical stimulus"
#| fig-name: "fig-sem_results_OB"
#
# A) Getting all the direct coefficients
OB_density_det <- paste0(format_dec(mean(OB_coeff_mitodensity_det), 3),
                  " [", format_dec(quantile(OB_coeff_mitodensity_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_mitodensity_det, 0.95), 3), "]")
OB_potential_det <- paste0(format_dec(mean(OB_coeff_potential_det), 3),
                  " [", format_dec(quantile(OB_coeff_potential_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_potential_det, 0.95), 3), "]")
OB_dna_det <- paste0(format_dec(mean(OB_coeff_dna_det), 3),
                  " [", format_dec(quantile(OB_coeff_dna_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_dna_det, 0.95), 3), "]")
OB_perox_det <- paste0(format_dec(mean(OB_coeff_perox_det), 3),
                  " [", format_dec(quantile(OB_coeff_perox_det, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_perox_det, 0.95), 3), "]")
#
OB_ros_dna <- paste0(format_dec(mean(OB_coeff_ros_dna), 3),
                  " [", format_dec(quantile(OB_coeff_ros_dna, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_ros_dna, 0.95), 3), "]")
#
OB_ros_perox <- paste0(format_dec(mean(OB_coeff_ros_perox), 3),
                  " [", format_dec(quantile(OB_coeff_ros_perox, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_ros_perox, 0.95), 3), "]")
#
OB_density_ros <- paste0(format_dec(mean(OB_coeff_mitodensity_ros), 3),
                  " [", format_dec(quantile(OB_coeff_mitodensity_ros, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_mitodensity_ros, 0.95), 3), "]")
OB_potential_ros <- paste0(format_dec(mean(OB_coeff_potential_ros), 3),
                  " [", format_dec(quantile(OB_coeff_potential_ros, 0.05), 3),
                  ", ", format_dec(quantile(OB_coeff_potential_ros, 0.95), 3), "]")
#
imgOB <- readPNG(here("Others", "SEM.png"))
plot_SEM_OB <- rasterGrob(imgOB, interpolate = TRUE)
#
fig_SEM_OB <- ggdraw(plot_SEM_OB) +
  draw_image("./Others/OB.png", x = 0.855, y = 0.67, width = 0.18, height = 0.3) +
  annotate("text", x = 0.52, y = 0.075, label = OB_density_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.52, y = 0.94, label = OB_potential_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.715, label = OB_dna_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.83, y = 0.335, label = OB_perox_det, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.525, y = 0.805, label = OB_ros_dna, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.515, y = 0.215, label = OB_ros_perox, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.282, y = 0.2, label = OB_density_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times") +
  annotate("text", x = 0.282, y = 0.82, label = OB_potential_ros, hjust = 1, vjust = 1, size = 2.8, family = "Times")
ggsave(here("./output/figures/text/SEM_OB.png"), plot = fig_SEM_OB, width = 21, height = 7.5, units = "cm", dpi = 600, bg = "white")
knitr::include_graphics("./output/figures/text/SEM_OB.png")
```

```{r, models_sem_OT}
#| label: models_sem_OT
# Making the same models above but for OT/Visual stimulus
#
source(here("R", "func.R"))
#
#
data_sem_OT <- SEM_df %>%
  filter(region == "OT") %>%
  mutate(obs = as.integer(c(1:80))) %>%
  mutate(vec = rep(1, length(obs)))
#
# Create the models
#
refit <- FALSE
#
if(refit){
  m_OT <- brm(
    bf(t_D | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + mean_dnadamage + mean_peroxidation + (1|clutch) + (1|p|obs)) +
    bf(mean_dnadamage | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch) + (1|q|obs)) +
    bf(mean_peroxidation | se(vec, sigma = FALSE) ~ mean_ros + (1|clutch) + (1|q|obs)) +
    bf(mean_ros | se(vec, sigma = FALSE) ~ mean_mitodensity + mean_potential + (1|clutch) + (1|t|obs)) +
  set_rescor(FALSE),
  family = gaussian(),
  data = data_sem_OT, 
  chains = 4, cores = 4, iter = 8000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth = 11))
  # Save the model
  saveRDS(m_OT, file = here("output/m_SEM/m_OT.rds"))
} else {
  m_OT <- readRDS(here("output/m_SEM/m_OT.rds"))
}
```

```{r, sem_tidy_OT}
#| label: sem_tidy_OT

@@

post_sem_OT <- as_draws_df(m_OT) 

@@

@@

OT_coeff_mitodensity_det <- post_sem_OT$b_tD_mean_mitodensity
OT_coeff_potential_det <- post_sem_OT$b_tD_mean_potential
OT_coeff_dna_det <- post_sem_OT$b_tD_mean_dnadamage
OT_coeff_perox_det <- post_sem_OT$b_tD_mean_peroxidation

@@

OT_coeff_ros_dna <- post_sem_OT$b_meandnadamage_mean_ros

@@

OT_coeff_ros_perox <- post_sem_OT$b_meanperoxidation_mean_ros

@@

@@

@@

@@

@@

  reference_variable = rep("t_D", 5),

@@

  direct_effects = I(list(OT_coeff_mitodensity_det,

@@

  indirect_effects = I(list(OT_undir_mitodensity_det,

@@

  total_effects = I(list(OT_total_mitodensity_det,

@@

  reference_variable = rep("DNA damage", 3),

@@

@@

  indirect_effects = I(list(OT_undir_mitodensity_dna,

@@

  total_effects = I(list(OT_total_mitodensity_dna,

@@

  reference_variable = rep("Lipid peroxidation", 3),

@@

@@

  indirect_effects = I(list(OT_undir_mitodensity_perox,

@@

  total_effects = I(list(OT_total_mitodensity_perox,

@@

#### Results of the SEM

Table S5. Compiled direct, indirect, and total effects of SEM for OB/Chemical stimulus
```{r, table_sem_results_OB}
#| label: table_sem_results_OB
#| fig-cap: ""
#
# Table created from the df sem_results_OB (see above)
source(here("R", "func.R"))
#
# Modify database
table_semOB_df <- sem_results_OB %>%
  mutate(`Direct effects` = paste0(mean_direct_effects, " [", q5_direct, ", ", q95_direct, "]"),
         `Indirect effects` = paste0(mean_indirect_effects, " [", q5_indirect, ", ", q95_indirect, "]"),
         `Total effects` = paste0(mean_total_effects, " [", q5_total, ", ", q95_total, "]")) %>%
  dplyr::select(source, predictor_modulator, `Direct effects`, `Indirect effects`, `Total effects`) %>%
  mutate(
      predictor_modulator = case_when(
        predictor_modulator == "mean_mitodensity" ~ "Mitochondrial density",
        predictor_modulator == "mean_potential" ~ "Mitochondrial potential",
        predictor_modulator == "mean_dnadamage" ~ "DNA damage",
        predictor_modulator == "mean_peroxidation" ~ "Lipid peroxidation",
        predictor_modulator == "ROS" ~ "ROS",
        TRUE ~ predictor_modulator
      ),
      source = case_when(
        source == "det_results_semOB" ~ "Detection latency",
        source == "dna_results_semOB" ~ "DNA damage",
        source == "perox_results_semOB" ~ "Lipid peroxidation",
        source == "ros_results_semOB" ~ "ROS",
        TRUE ~ source
      ),
      `Direct effects` = case_when(`Direct effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Direct effects`),
      `Indirect effects` = case_when(`Indirect effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Indirect effects`),
      `Total effects` = case_when(`Total effects` == "NaN [NA, NA]" ~ " - ", TRUE ~ `Total effects`)) %>%
  rename(Predictor = predictor_modulator, Response = source) %>%
  mutate(Response = factor(Response, levels = c("Detection latency", "DNA damage", "Lipid peroxidation", "ROS")),
         Predictor = factor(Predictor, levels = c("Mitochondrial density",
                                                "Mitochondrial potential",
                                                "ROS",
                                                "DNA damage",
                                                "Lipid peroxidation"))) %>%
  arrange(Response, Predictor)
#
# Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
#
table_semOB <- flextable(table_semOB_df) %>%
  align(align = "center", part = "header") %>%
  flextable::compose(i = c(2:5,7,8,10,11,13), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = c(5,8,11), part = "body") %>% 
  autofit()
#
table_semOB
```
